---
title: "hw1"
author: "shuangyu_zhao"
date: "2023-01-11"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(sqldf)
```

```{r}
order_details <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/order_details.csv")
orders <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/orders.csv")
territories <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/territories.csv")
regions <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/regions.csv")
employee_territories <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/employee_territories.csv")
employees <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/employees.csv")
customers <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/customers.csv")
shippers <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/shippers.csv")
suppliers <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/suppliers.csv")
products <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/products.csv")
categories <- read.csv("/Users/apple/Desktop/STT811 appl_stat_model/data/northwind/categories.csv")
```

1.Perform a sort of orders by employeeID, then by shipVia, and then by freight, for those orders by shipped to France.
```{r}
ordered_order <- sqldf("SELECT *
                        FROM orders
                        WHERE shipCountry = 'France'
                        ORDER BY employeeID, shipVia, freight")
head(ordered_order)
```



2. Which shipVia has the largest average cost?
```{r}
ave_cost_max <- sqldf("SELECT shipVia, 
                  AVG(freight) AS ave_cost
                  FROM orders
                  GROUP BY shipVia
                  ORDER BY ave_cost DESC
                  LIMIT 2")
ave_cost_max
```

shipVia = 2 has the largest average cost.


3. Which product category has the highest average UnitPrice? The Lowest?
```{r}
categorie_unitprice<- sqldf("SELECT AVG(UnitPrice), categories.*
                             FROM products
                             INNER JOIN categories 
                             ON categories.categoryID = products.categoryID
                             GROUP BY products.categoryID
                             ORDER BY AVG(UnitPrice)")
head(categorie_unitprice, 2)
tail(categorie_unitprice, 2)
```

Meat/Poultry(6) has the highest average UnitPrice.

Grains/Cereals(5) has the lowest average UnitPrice.



4. Which products are supplied by a company in the United States?
```{r}
product_usa <- sqldf("SELECT Country, CompanyName, suppliers.SupplierID, products.ProductID, products.ProductName
                    FROM suppliers
                    INNER JOIN products 
                    ON suppliers.SupplierID = products.SupplierID
                    GROUP BY products.ProductID
                    HAVING Country = 'USA'
                    ORDER BY suppliers.SupplierID")
product_usa
```

Chef Anton's Cajun Seasoning, Chef Anton's Gumbo Mix, Louisiana Fiery Hot Pepper Sauce and Louisiana Hot Spiced Okra are supplied by New Orleans Cajun Delights in USA.

Grandma's Boysenberry Spread, Uncle Bob's Organic Dried Pears and Northwoods Cranberry Sauce are supplied by Grandma Kelly's Homestead in USA.

Sasquatch Ale, Steeleye Stout and Laughing Lumberjack Lager are supplied by Bigfoot Breweries in USA.

Boston Crab Meat and Jack's New England Clam Chowder are supplied by New England Seafood Cannery in USA.




5.Which shipper is shipping the largest number of units of product? Answer in terms of units; you do not need to consider quantityPerUnit here.
```{r}
shipper_largest <- sqldf("SELECT order_details.quantity, 
                         orders.shipName, orders.shipVia,
                         shippers.companyName
                         FROM order_details
                         INNER JOIN orders ON orders.orderID = order_details.orderID
                         INNER JOIN shippers ON orders.shipVia = shippers.shipperID
                         ORDER BY order_details.quantity DESC
                         LIMIT 3")
shipper_largest
```

Federal Shipping and United Package are shipping the largest number of units of products.



6. Which employee is tied to the most sales revenue? Give the name, not the code, along with the total revenue for the employee.
```{r}
order_revenue <- sqldf("SELECT orders.orderID, orders.employeeID,
                       order_details.unitPrice * order_details.quantity *( 1- order_details.discount) AS revenue
                       FROM order_details
                       INNER JOIN orders
                       WHERE order_details.orderID = orders.orderID")

best_employee <- sqldf("SELECT revenue, order_revenue.employeeID, 
                       employees.firstName, employees.lastName
                       FROM order_revenue
                       INNER JOIN employees
                       WHERE employees.employeeID = order_revenue.employeeID
                       ORDER BY revenue DESC
                       LIMIT 2")
best_employee
```
Nancy Davolio is tied to the most sales revenue.


7. Find the total revenue for each product category.
```{r}
product_order_revenue <- sqldf("SELECT productID, orderID,
                              order_details.unitPrice * order_details.quantity *( 1- order_details.discount) AS revenue
                               FROM order_details
                              ORDER BY orderID")

totallrevenue_category <- sqldf("SELECT SUM(product_order_revenue.revenue) AS revenue_category,
                                products.CategoryID, categories.categoryName
                                FROM product_order_revenue
                                INNER JOIN products ON products.productID = product_order_revenue.productID
                                INNER JOIN categories ON products.CategoryID = categories.categoryID
                                GROUP BY products.CategoryID")
totallrevenue_category
```



8. Consider the amount of revenue for each customer. If there were no discounts applied, which customer would see the largest increase in cost?
```{r}
order_revenue_incre <- sqldf("SELECT orders.orderID, orders.customerID, customers.companyName,
                       order_details.unitPrice * order_details.quantity * order_details.discount AS revenue_increase
                       FROM orders
                       INNER JOIN order_details ON order_details.orderID = orders.orderID
                       INNER JOIN customers ON orders.customerID = customers.customerID
                       ORDER BY revenue_increase DESC")
head(order_revenue_incre)
```

Piccolo und mehr and Queen Cozinha have largest increase in cost. 


9. Which order(s) has the most number of items (and how many)? Give the orderID for this one.
```{r}
order_item <- sqldf("SELECT order_details.orderID,
                    SUM(order_details.quantity) AS item_num
                    FROM order_details
                    GROUP BY order_details.orderID
                    ORDER BY item_num DESC
                    ")

head(order_item)

```

10895 has the most items.

10. Create a new field called “InventoryOrderRatio” which is, for each product, the UnitsinStock (the inventory) for the product (across all customers) divided by the quantity ordered for that product. A high value represents sufficient product in stock, while a low number represents products that are in danger of running out. What 3 products are most in danger of running out?
```{r}
products_ior <- sqldf("SELECT CAST(products.UnitsInStock AS FLOAT) / CAST(SUM(order_details.quantity) AS FLOAT) AS InventoryOrderRatio, 
                      products.ProductID, products.ProductName
                      FROM products
                      INNER JOIN order_details ON order_details.productID = products.productID
                      GROUP BY order_details.productID
                      ORDER BY InventoryOrderRatio")
head(products_ior)
```
Chef Anton's Gumbo Mix, Alice Mutton, Thüringer Rostbratwurst are most in danger of running out


11. A recommender engine looks at which pairs of products tend to be bought by the same customer, so that if a customer buys one, the recommender engine will recommend they buy the other. Find which product pairs are most likely to be bought by the same customer.
```{r}
orders_sim <- sqldf("SELECT orders.orderID, orders.customerID, order_details.productID
               FROM orders
               INNER JOIN order_details 
               ON orders.orderID = order_details.orderID")
head(orders_sim)
```
```{r}
order_times <- sqldf("SELECT customerID, COUNT( DISTINCT orderID) AS order_time
                      FROM orders_sim 
                      GROUP BY customerID")
head(order_times)
```


```{r} 
pairs_order <-  sqldf("SELECT tabel1.customerID, tabel1.productID AS item1, tabel2.productID AS item2
                      FROM orders_sim AS tabel1
                      INNER JOIN orders_sim AS tabel2
                      ON tabel1.orderID = tabel2.orderID
                      AND tabel1.productID < tabel2.productID
                      GROUP BY tabel2.customerID, tabel2.orderID, tabel1.productID, tabel2.productID")
head(pairs_order, 20)
```
```{r}
pair_order_times <- sqldf("SELECT pairs_order.*, COUNT(*) AS times
                          FROM pairs_order
                          GROUP BY customerID, item1, item2
                          ORDER BY times DESC")
head(pair_order_times)
```

```{r}
prob_pair <- sqldf("SELECT pair_order_times.*, order_times.order_time AS total_order_times, 
                   100 * (CAST(pair_order_times.times AS FLOAT) / CAST(order_times.order_time AS FLOAT)) AS prob_percentage
                   FROM pair_order_times
                   INNER JOIN order_times
                   ON pair_order_times.customerID = order_times.customerID
                   ORDER BY prob_percentage DESC")
head(prob_pair)
```
21 and 37 are more likely to be bought by CENTC


