---
title: "ica19_syz"
author: "Shuangyu Zhao"
date: "2023-03-28"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.
```{r}
passenger <- read.csv("/Users/apple/Desktop/STT811_appl_stat_model/data/passenger.csv")
head(passenger, 13)
```

 b.	Convert it into a time series
```{r}
passenger_ts <- ts(passenger$AirPassengers, frequency = 12, start = c(1949, 1), end = c(1960,12))
```

c.	Create the autocorrelation and partial autocorrelation panels.  Comment on how the seasonality is observed
```{r}
acf(passenger_ts)
pacf(passenger_ts)

```

d.	Perform a decomposition of the time series, trying both additive and multiplicative.  Which one “looks” better?
```{r}
pass_multi <- decompose(passenger_ts, type = "multiplicative")
plot(pass_multi)
```

```{r}
pass_add <- decompose(passenger_ts, type = "additive")
plot(pass_add)
```
additive looks better

2.
```{r}
oil_gas <- read.csv("/Users/apple/Desktop/STT811_appl_stat_model/data/oil-gas.csv")
head(oil_gas)
tail(oil_gas)
```

b.	Convert the oil and gas data into 2 separate time series (what is the seasonality frequency?)
```{r}
oil_ts <- ts(oil_gas$Oil, frequency = 52, start = c(2013, 1), end = c(2022,4))
gas_ts <- ts(oil_gas$Gas, frequency = 52, start = c(2013, 1), end = c(2022,4))
```


c.	Perform a decomposition on the oil data.  How strong is the seasonality?
```{r}
oil_multi <- decompose(oil_ts, type = "multiplicative")
plot(oil_multi)
oil_add <- decompose(oil_ts, type = "additive")
plot(oil_add)
```


d.	Create a scatterplot of gas vs. oil (oil on x axis).  How correlated are they?
```{r}
plot(oil_gas$Oil, oil_gas$Gas)
```
they have great correlation










